Stateless Model Checking with Data-Race Preemption Points

### 2016-03-15 oopsla-16/paper.tex
start doing revisions for oopsla

% template-provided usepackages
\usepackage{amsmath}
% custom ones copied from pldi template
\usepackage{url}
\usepackage{enumitem}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}

@@

% ???
%\newcommand{\cL}{{\cal L}}

@@

@@

We present an {\em Iterative Deepening} framework for stateless model checking,
which manages the exploration of many state spaces using different preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates data-race analysis to add new preemption points on the fly.
Preempting threads during a data race's instructions
%These preemption points allow us to check whether each data-race can produce an observable failure, and
can automatically classify the race as buggy or benign,
and uncovers new bugs not reachable by prior model checkers.
%
It also enables full verification of all possible schedules when every data race is verified as benign within the CPU budget.
In our evaluation, Iterative Deepening
%with data-race preemption points
found 1.8x as many bugs and verified 4.3x as many tests compared to stateless model checking without data-race preemption points.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.
\end{abstract}

\category{D.2.4}{Software Engineering}{Software/Program Verification}
\category{D.2.5}{Software Engineering}{Testing and Debugging}
%
%% general terms are not compulsory anymore,
%% you may leave them out
%\terms
%term1, term2
%
%\keywords
%keyword1, keyword2

>>>>

\usepackage{amsmath}
\usepackage{url}
\usepackage{enumitem}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{courier} % TODO: remove this? and get nicer fonts?
\usepackage{amsthm}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}

@@

@@

% TODO: remove all changebars before submission
%\setcounter{changebargrey}{0}
%\newcommand\revision[1]{\begin{samepage}\cbstart#1\cbend\end{samepage}}
%\newcommand\revision[1]{\cbstart#1\cbend}
%\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
\newcommand\revision[1]{\hilight{purple}{#1}}

@@

\revision{We present \quicksand, a new stateless model checking framework}~which
manages the exploration of many state spaces using different preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates data-race analysis to add new preemption points on the fly.
Preempting threads during a data race's instructions
%These preemption points allow us to check whether each data-race can produce an observable failure, and
can automatically classify the race as buggy or benign,
and uncovers new bugs not reachable by prior model checkers.
%
It also enables full verification of all possible schedules when every data race is verified as benign within the CPU budget.
In our evaluation, Iterative Deepening
%with data-race preemption points
found 1.8x as many bugs and verified 4.3x as many tests compared to stateless model checking without data-race preemption points.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.
\end{abstract}

\revision{\category{D.2.4}{Software Engineering}{Software/Program Verification}}
%\category{D.2.5}{Software Engineering}{Testing and Debugging}
%
%% general terms are not compulsory anymore,
%% you may leave them out
%\terms
%term1, term2
%
%\keywords
%keyword1, keyword2

### 2016-03-15 oopsla-16/related.tex
start doing revisions for oopsla

{\bf Synchronization events only.} CHESS \cite{chess} and dBug \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence, they will miss any bugs that require interleaving threads at instruction granularity during a data race. CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data-race candidates as PPs in future tests.

@@

{\bf Other techniques.} Various improvements or alternatives to DPOR have been developed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr},
%DPOR for relaxed memory models \cite{tsopso},
and SAT-directed MC \cite{satcheck}.
These are all compatible with our technique.
Recent work \cite{tsopso} has extended DPOR for relaxed memory models \cite{memory-consistency-models},
which we do not yet account for in our proofs (\sect{\ref{sec:soundness}}).
Parrot \cite{parrot} combines MC with a partially-determinizing runtime for further reduction, but still, fewer than half the non-trivial state spaces in their evaluation could be completed.
%providing a strong argument for \quicksand.
Finally, Iterative Context Bounding (ICB) \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs, and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan} combine the lockset and happens-before analyses into a hybrid technique, which we employ.
DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%event-driven
Android applications, using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} offers data-race techniques for kernel code.
IFRit \cite{ifrit}
improves performance using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-

>>>>

{\bf Synchronization events only.} CHESS \cite{chess} and dBug \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence, they will miss any bugs that require interleaving threads at instruction granularity during a data race.
%CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data-race candidates as PPs in future tests.
\revision{\cite{chess} mentions the ability of CHESS to add PPs using a single-pass data-race analysis,
% FIXME... idk, "discusses (without explicit evaluation)..., but, as our evaluation shows..."?
but does not discuss either the increase in bug-finding capacity or the soundness properties.
Our convergence theorem (\sect{\ref{sec:soundness}}) is a natural extension of \cite{chess-icb}'s Theorem 3,
which alone provides soundness only for race-free programs.}

@@

{\bf Other techniques.} Various improvements or alternatives to DPOR have been developed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr},
%DPOR for relaxed memory models \cite{tsopso},
SAT-directed MC \cite{satcheck},
\revision{and R4 for event-driven applications \cite{r4}}.
These are all compatible with our technique.
Recent work \cite{tsopso} has extended DPOR for relaxed memory models \cite{memory-consistency-models},
which we do not yet account for in our proofs (\sect{\ref{sec:soundness}}).
Parrot \cite{parrot} combines MC with a partially-determinizing runtime for further reduction, but still, fewer than half the non-trivial state spaces in their evaluation could be completed.
%providing a strong argument for \quicksand.
Finally, Iterative Context Bounding (ICB) \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs, and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan} combine the lockset and happens-before analyses into a hybrid technique, which we employ.
\revision{Several tools \cite{droidracer,cafa,unnamed-android} have recently emerged to target Android applications},
%DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%%event-driven
%Android applications,
using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} offers data-race techniques for kernel code.
IFRit \cite{ifrit}
improves performance using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-
